---
title: "Reproducible Research - Project 1"
output:
  html_document:
    keep_md: yes
---

Opens file, loads dplyr package
```{r}
data <- read.csv("C:/Users/Noah/Documents/Coursera/Reproducable research/project 1/activity.csv")
library(dplyr)
```

splits data by date, gets sum/mean/median no of steps
```{r}
dates <- group_by(data, date)
steps <- summarise(dates, meansteps=mean(steps, na.rm=TRUE),
                        totalsteps = sum(steps, na.rm=TRUE), 
                        mediansteps=median(steps, na.rm=TRUE))
steps
```
creates histogram
```{r}
hist(steps$totalsteps, breaks = 11, main = "Total steps/day",xaxt="n")
axis (side=1, at=seq(0,22000,2000), labels = seq(0,22000,2000))
```
splits data by interval,finds interval with highest avg
```{r}
interval <- group_by(data, interval)
stepsinterval <- summarise(interval, meansteps=mean(steps, na.rm=TRUE))
max <- which.max(stepsinterval$meansteps)
maxinterval<-stepsinterval[max,1]
print(paste0("the interval with highest average number of steps is ",maxinterval))
```
creates time series plot
```{r}
library(ggplot2)
ggplot(stepsinterval, aes(interval, meansteps)) + geom_line() + scale_x_continuous(breaks=seq(0,2500, 100))
```
find n/as in steps var
```{r}
NAvals <- is.na(data$steps)
table(NAvals)
```

find avg steps for each interval
```{r}
data2 <- data
data2$avg <- tapply(data2$steps, data2$interval, mean,na.rm=TRUE)
```

replace nas with avg steps
all values in newsteps var. that are N/As are replaced with the corresponding value in that row of the avg. variable
```{r}
data2$newsteps <- data2$steps
data2$newsteps[is.na(data2$newsteps)] = data2$avg[is.na(data2$newsteps)]
```

splits data (that now has no NAs) by date,gets sum/mean/median no of steps
```{r}
datesnoNA <- group_by(data2, date)
stepsnoNA <- summarise(datesnoNA, meansteps_noNA=mean(newsteps, na.rm=TRUE),
                   totalsteps_noNA = sum(newsteps, na.rm=TRUE), 
                   mediansteps_noNA=median(newsteps, na.rm=TRUE))
stepsnoNA
```
creates histogram
```{r}
hist(stepsnoNA$totalsteps, breaks = 11, main = "Total steps/day",xaxt="n")
axis (side=1, at=seq(0,22000,2000), labels = seq(0,22000,2000))
```

compares steps w/ and w/o NAs
```{r}
stepscompare <- cbind(steps, stepsnoNA[,2:4])
head(stepscompare)
```

formats dates as date format
if weekday is sat or sun, var is weekend, else its weekday
```{r}
data2$date <- as.Date(data2$date)
data2$weekdays <- weekdays(data2$date)
data2$day <- ifelse((data2$weekdays == "Saturday" | data2$weekdays == "Sunday"), "weekend", "weekday")
```
converts to factor
```{r}
data2$day <- as.factor(data2$day)
```

creates time plot
```{r}
weekend_day <- group_by(data2, interval,day)
stepsintervalday <- summarise(weekend_day, meansteps=mean(newsteps, na.rm=TRUE))
graph <-ggplot(stepsintervalday, aes(interval, meansteps)) + 
    geom_line() + scale_x_continuous(breaks=seq(0,2500, 100)) 
graph + facet_wrap(~day, ncol=1)
```
